Collocated Deduplication with Fault Isolation for Virtual Machine Snapshot Backup

### 2013-07-26 snapshot_system/arch.tex
update deletion

In a busy VM cluster, snapshot deletions are as frequent as snapshot creations.
The VM-centric snapshot storage design simplifies the deletion process since we need to locate unreferenced chunks within each VM to free up space when deleting a snapshot.
The PDS data chunks are commonly shared all VMs and we do not consider their reference
counting  during snapshot deletion.
The selection of PDS data chunks   is updated periodically independent of snapshot deletion process.

While we can use the standard mark-sweep technique~\cite{mark-sweep}, 
it takes significant time to conduct this process every time there is a snapshot deletion.
In the case of Alibaba, snapshot backup is conducted automatically and there are 
about 10 snapshot stored for  every user. Then when there is
a new snapshot created every day,  there will be  a snapshot expired everyday to maintain
a balanced storage use. Given the large number of snapshot deletion requests, we seek
a fast solution with a very low resource usage to delete snapshots.

@@

%Our system adopts VM-centric snapshot  lazy delete strategy so that all snapshot deletions are scheduled
%in the backup time window at midnight. 
%Therefore, snapshot deletions must be fast enough to fit in time window and
%efficient enough to satisfy our resource constraints.
%However, there is no simple solution can achieve these goals with high reliability.
%Our hybrid deletion strategy, using fuzzy deletion regularly and accurate deletion periodically,
%accomplishs our speed, resource usage and relibility goals very well.
\begin{itemize}
\item {\bf Computation for snapshot fingerprint summary.}
Every time there is a new snapshot created,
we compute a bloom-filter fingerprint summary for all non-PDS chunks used 
in this snapshot.
%{\bf Creating bloom filter} Scan all the living snapshot recipess and their segment recipes,
%for every reference pointing to append store, add it to the bloom filter.

@@

\item {\bf Periodic repair of leakage}
%[exlpain why second bloom filter, why scan append store]
Since there are certain unused chunks which are not deleted during
the approximate deletion, leakage repair is conducted periodically.
In this phase we load the metadata of all containers from a VM snapshot store,
and mark all used chunks. Unmarked (and therefore unused) chunks within this VM can be safely deleted.
This process is identical to the previously-developed mark-and-sweep techniques except that
we only need to conduct a VM-specific mark-and-sweep  operation, and only do that periodically.
%We cannot simply repeat the phase 1 multiple times to reduce temporary storage leakage, because:
%\begin{enumerate} 
%\item After several runs of phase-1, it is proven that the merged type-1 summaries cannot sieve remaining unused blocks, due to the false-positive property of bloom filter.
%\item The recipes of deleted snapshots have been removed from the system, thus we are not able to obtain the deleted block fingerprints from any metadata, the only way to discover them is to scan the append store indices.
%\end{enumerate}
\end{itemize}

@@

%{\bf Discussion}
We estimate how often the leak repair needs to be conducted.
Let $L$ be the amount of temporary storage leakage caused by our approximate snapshot deletion, 
$H$ be the average size of non-PDS snapshot data that should be deleted. 
$\epsilon$ be the false-positive rate of the merged bloom filter. 
$R$ be the number of runs using approximate deletion before a full leak repair is conducted.
Then $ L= R * \epsilon * H.  $

When ratio $L/H$ exceed a threshold $\tau$, then:
\[
\frac{L}{H} = R * \epsilon > \tau \Rightarrow R > \frac{\tau}{\epsilon}.
\]

For example, when $\epsilon = 0.01$ and $\tau=3$, 
we would expect deletion phase-2 be triggered once for 
every $T/P_{bl} = 300$ runs of fast deletion. 
When one machine hosts 25 VMs and there is one snapshot deletion per day per VM, there would be 
only one full leak repair scheduled for every 12 days, which is sufficiently small to offset the 
heavy I/O workload of the mark-sweep process.

>>>>

In a busy VM cluster, snapshot deletions are as frequent as snapshot creations.
However, data deduplication complicates the deletion process because space saving relies on the sharing of data,
thus making it difficult to decide which chunks are safe to delete.
In a traditional deduplication system, deletion often require looking at global scope to
resolve the data dependency between logical backup entities and physical data chunks.
Our VM-centric snapshot storage design simplifies the deletion process since 
we only need to locate unreferenced chunks within each VM's snapshot store to free up space when deleting a snapshot.
The PDS data chunks are commonly shared all VMs and we do not consider their reference
counting during snapshot deletion.
%The selection of PDS data chunks is updated periodically independent of snapshot deletion process.

While we can use the standard mark-and-sweep technique~\cite{mark-sweep}, 
it still takes significant time to conduct this process every time there is a snapshot deletion.
In the case of Alibaba, snapshot backup is conducted automatically and there are 
about 10 snapshot stored for every user. When there is
a new snapshot created every day,  there will be  a snapshot expired everyday to maintain
a balanced storage use. Given the large number of snapshot deletion requests, we seek
a fast solution with a very low resource usage to delete snapshots.

@@

%Our system adopts VM-centric snapshot  lazy delete strategy so that all snapshot deletions are scheduled
%in the backup time window at midnight. 
%Therefore, snapshot deletions must be fast enough to fit in time window and
%efficient enough to satisfy our resource constraints.
%However, there is no simple solution can achieve these goals with high reliability.
%Our hybrid deletion strategy, using fuzzy deletion regularly and accurate deletion periodically,
%accomplishs our speed, resource usage and relibility goals very well.
\begin{itemize}
\item {\bf Computation for snapshot fingerprint summary.}
Every time there is a new snapshot created,
we compute a bloom-filter as the summary of reference pointers for all non-PDS chunks used 
in this snapshot.
%{\bf Creating bloom filter} Scan all the living snapshot recipess and their segment recipes,
%for every reference pointing to append store, add it to the bloom filter.

@@

\item {\bf Periodic repair of leakage}
%[exlpain why second bloom filter, why scan append store]
Since there are certain unused chunks which are not deleted during
the approximate deletion, leakage repair is conducted periodically.
In this phase we load the reference pointers of all chunks from a VM snapshot store as a table in memory,
then scan through all the snapshots' segment recipes. For each reference pointer that is in use by a snapshot,
we mark the corresponding entry in the table as in use. Finally those unmarked (and therefore unused) reference pointers in the
table represent chunks that can be safely deleted.

The cost of leak repair mainly come from holding a table of reference pointers and scan of all snapshots metadata.
Consider each reference pointer consumes 8 bytes plus 1 byte as mark field, a VM that has 40GB backup data will
need 90MB of memory to construct such a table. Also, scanning all snapshots metadata is many times slower comparing
to the approximate deletion which only scans single snapshot's metadata.
It's worth to mention that compare to previously-developed mark-and-sweep techniques, 
our leak repair still has advantage because the scope of repair is restricted within single VM's snapshot 
store due to our VM-specific design, and we only do it periodically.
%We cannot simply repeat the phase 1 multiple times to reduce temporary storage leakage, because:
%\begin{enumerate} 
%\item After several runs of phase-1, it is proven that the merged type-1 summaries cannot sieve remaining unused blocks, due to the false-positive property of bloom filter.
%\item The recipes of deleted snapshots have been removed from the system, thus we are not able to obtain the deleted block fingerprints from any metadata, the only way to discover them is to scan the append store indices.
%\end{enumerate}
\end{itemize}

@@

%{\bf Discussion}
We now give a model to estimate the storage leakage and how often leak repair needs to be conducted.
Assuming a VM always maintain $B$ snapshots in the backup, and it creates and deletes one snapshot
everyday. Let $c$ be the total number of chunks brought by the initial backup, $c^'$ be the average
number of chunks brought by a consequancial backup, then the total number of chunks in use $C$ is:
\[
C = c + (B-1)c^'
\]

Let $m$ be the bloom filter size in bits, $k$ be the number of hash functions used by bloom filter,
then bloom filter's false-positive rate $\epsilon$ is:
\[
\epsilon = (1-e^{\frac{kC}{m}})^k
\]

For each snapshot deletion, the amount of chunks need to be deleted is nearly identical to the number of
newly add chunks $c^'$. However, some of chunks among them are not detected as unused in our approximation
algorithm, thus forms the storage leakage. Let $R$ be the total number of runs of approximate deletion since
last repair, the total leakage $L$ can be estimated as:
\[
L = Rc^'\epsilon
\]

When leakage ratio $L/C$ exceeds a pre-defined threshold $\tau$, we need to execute a repare:
\[
\frac{L}{C} = \frac{Rc^'\epsilon}{c+(B-1)c^'} > \tau \rightarrow R > \frac{\tau}{\epsilon}\frac{c + (B-1)c^'}{c^'}
\]
For example, when $\epsilon = 0.01$ and $\tau=0.1$, when the VM keeps 10 snapshots and each snapshot contains
about 5\% of new data, then
we would expect leak repair be triggered once for 
every 290 runs of approximate deletion. 
When one machine hosts 25 VMs and there is one snapshot deletion per day per VM, there would be 
only one full leak repair scheduled for every 12 days, which is sufficiently small to offset the 
heavy I/O workload of the mark-sweep process.

### 2013-07-26 snapshot_system/snapshot.tex
update deletion

Since the best deduplication uses a nonuniform chunk size 
in the average of 4K or 8K~\cite{??},  
%This allows the system to achieve deduplication even when there are insertions/deletions which would
%affect many fixed-size chunks.
we conduct additional local inner-VM deduplication by comparing
chunk signatures within a dirty segment from its parent snapshot. 
We load the chunk fingerprints in the corresponding segment from the
parent and perform fingerprint matching for further inner-VM deduplication.
The amount of memory for maintaining those fingerprints  is small, as we only
load one segment at a time.
For example, with a 2MB segment, there are about 500 fingerprints to compare.

>>>>

Since the best deduplication uses a nonuniform chunk size 
in the average of 4K or 8K~\cite{??},  
%This allows the system to achieve deduplication even when there are insertions/deletions which would
%affect many fixed-size chunks.
we conduct additional local inner-VM deduplication by comparing
dirty segment's chunk fingerprints to similar segments from its parent snapshot. 
For every segment, the minimum value of all its chunk fingerprints is computed during backup
and is recorded in the snapshot recipe.
So when we are processing a dirty segment,
we can easily find similar segments from the
parent snapshot recipe and load their segment recipes. Then we can perform fingerprint matching 
against them
for further inner-VM deduplication.
The amount of memory for maintaining those fingerprints is small, as we only
load one segment at a time.
For example, with a 2MB segment, there are about 500 fingerprints to compare.
